---
title: "Oxygen over time : TEOS Hydrostation S Data Lab"
author: "Jose Martinez"
date: "2023-03-16"
output: html_document
---
# Load required libraries
```{r setup, include=FALSE}
library(prettydoc)
library(tidyverse)
library(gsw)
library(readr)
library(plotly)
```

## Now we need to import our data
```{r message = FALSE, warning = FALSE}
hydrostation_bottle <- read_delim("hydrostation_bottle.txt", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE, skip = 31)

hydrostation_bottle_names <- read_csv("hydrostation_bottle.txt", 
    skip = 30)

colnames(hydrostation_bottle) = colnames(hydrostation_bottle_names)
```
## Hydrostation S Discrete Bottle Data for years 1955 through December 2020.

### Variable Names and Units

- yyyymmdd = Year Month Day   
- decy   = Decimal Year     
- time   = Time (hhmm)      
- latN   = Latitude (Deg N) 
- lonW   = Longitude (Deg W)
- Depth  = Depth (m)                  
-Temp   = Temperature ITS-90 (C) 
- Pres   = CTD Pressure (dbar)   
- CTD_S  = CTD Salinity (PSS-78)      
- Sal1   = Salinity-1 (PSS-78)        
- Sig-th = Sigma-Theta (kg/m^3)       
- O2(1)  = Oxygen-1 (umol/kg)          
- OxFixT = Oxygen Fix Temp (C)        
- Anom1  = Oxy Anomaly-1 (umol/kg)    
Quality flags
- -999  = No data
- 0 = Less than detection limit
```{r}
hydroS =
  hydrostation_bottle %>% 
  filter(Sal1!=-999) %>%
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1,latN,)) %>% 
  mutate(S_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN)) %>% 
  mutate(T_cons_gsw = gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw)) %>% 
  mutate(Sig_th_gsw = gsw_sigma0(S_abs_gsw,T_cons_gsw))

# how to replace data in a line without making this many dataframes (i.e. select and/or filters)
hydroS_correctedS_a =
  hydroS %>% 
  filter(Sig_th_gsw<0) %>%
  mutate(S_abs_gsw=gsw_SA_from_SP(CTD_S,Pres_gsw,360-lonW,latN)) %>% 
  mutate(T_cons_gsw = gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw)) %>% 
  mutate(Sig_th_gsw = gsw_sigma0(S_abs_gsw,T_cons_gsw)) 


hydroS_correctedS_b =
  hydroS %>% 
  filter(Sig_th_gsw>0)

hydroS_corrected = rbind(hydroS_correctedS_a, hydroS_correctedS_b) 
```
#Lab Assignment
1. Pick a question (include hypothesis)

How does oxygen behave over time? 
H1 : Oxygen reaches a min in the upper 1000m before increasing with further depth

2. Produce a plot and statistical summary using lm()
```{r}
o2_ht = 
  hydroS_corrected %>% 
  filter(`O2(1)`!= -999) %>% 
  filter(Depth < 50)

plot_o2vt = o2_ht %>% 
  ggplot()+
  geom_point(aes(x=decy,y = `O2(1)`))+
  ylab('Oxygen (µmol/kg)')+
  xlab('Time')+
  geom_smooth(aes(x=decy, y=`O2(1)`),method = "lm")

plot_o2vt

oxygen_time = lm(`O2(1)`~decy,data =  o2_ht) # y (Sig) is a function of x(decy), lm(y~x,data = data )
summary(oxygen_time)
```

## Lab Assignment


3. Describe your results, the summary, and answer the question

Oxygen ranges from a low of 155.8 umol/kg to a maximum of 270.5 umol/kg over the late 1960s to present day. We see higher oxygen values from the mid 1980s to the mid 1990s which can be an indicator of a positive NAO in contrast to the probable negative NAO trend that followed the mid 1990s into present. In the summary statisitics we see a slight negative slope of oxygen over time that is not statistically significant but is a probable indicator of anthropogenic climate change lowering oxygen content in the North Atlantic.

4. Compile into a completed lab report using R markdown
